Development of restored dike grasslands to reference needs more than four years and can hardly modified by local site conditions

### 2023-02-07 manuscript.Rmd
Update manuscript.Rmd

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)); 314 m above sea level; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table @ref(tab:flood-weather), @lfu2021waterlevel; @lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from removed agricultural soil of a dike construction site near Steinkirchen. The substrates were mixed with a big roller and an excavator prepared the plots. The target vegetation were lowland hay meadows [EUNIS code: R22, @Chytry.2020; Arrhenatherion elatioris, EuroVegChecklist: CM01A, @mucina2016] and semi-dry calcareous grassland (R1A; Cirsio-Brachypodion pinnati, DA01B). The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits))<!--# describe difference --> . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition, halfway up the slopes of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three sand admixtures (0%, 25% and 50%) and tow soil depths (15 vs. 30 cm) of practical interest [@KleberLerchbaumer.2017] and two . Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. We tested soil samples of the three substrates on both expositions by mixing several sub-samples. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table @ref(tab:substrate)). The two seed mixture types were tested with the normal seeding density and the increased density recommended for slopes [4 vs. 8 g m^-2^ @Kiehl.2010; @KleberLerchbaumer.2017]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species (Appendix Table \@ref(tab:establishment)).

@@

```{r tab.id = "substrate", ft.align = "left",tab.cap = "Soil of the three substrates was analyzed for the fraction < 2mm fraction. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl2 solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl2 extract. For calculating CaCO3, a subsample was annealed at 550 °C and the measured C amount calculated with 8.33. To calculate total N and the C/N ratio, a subsample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand."}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```

>>>>

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)); 314 m above sea level; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table @ref(tab:flood-weather), @lfu2021waterlevel; @lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from removed agricultural soil of a dike construction site near Steinkirchen. The substrates were mixed with a big roller and an excavator prepared the plots. The target vegetation were lowland hay meadows [EUNIS code: R22, @Chytry.2020; Arrhenatherion elatioris, EuroVegChecklist: CM01A, @mucina2016] and semi-dry calcareous grassland (R1A; Cirsio-Brachypodion pinnati, DA01B). The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits)). As expected have the hay meadows mixtures a higher specific leaf area (SLA) than the dry meadows mixtures<!--# describe difference --> . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition, halfway up the slopes of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three sand admixtures (0%, 25% and 50%) and tow soil depths (15 vs. 30 cm) of practical interest [@KleberLerchbaumer.2017] and two . Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. We tested soil samples of the three substrates on both expositions by mixing several sub-samples. The sand admixture changed the soil texture, increased the C/N ratio, but reduced Calciumcarbonat, and hardly changed the rather high pH values (Table\@ref(tab:substrate)). Phosphate and potassium were rather scarce for agricultural soils and Magnesium rather high [@lfl2022]. The two seed mixture types were tested with the normal seeding density and the increased density recommended for slopes [4 vs. 8 g m^-2^ @Kiehl.2010; @KleberLerchbaumer.2017]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species (Appendix Table \@ref(tab:establishment)).

@@

```{r tab.id = "substrate", ft.align = "left",tab.cap = "Soil of the three substrates was analyzed for the fraction < 2mm fraction. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl2 solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl2 extract. For calculating CaCO3, a subsample was annealed at 550 °C and the measured C amount calculated with 8.33. To calculate total N and the C/N ratio, a subsample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand."}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  select(-starts_with("lfl")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      select(-starts_with("lfl")) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```
